---
title: "Working with RSQLite"
author: "Kenneth Benoit (w/inspiration from **RSQLite** vignette)"
output: github_document
---
    
```{r, echo = FALSE}
knitr::opts_chunk$set(comment = "##", collapse = TRUE)
```

## Why and how to use RSQLite

- RSQLite is the easiest way to use a database from R because the package itself contains [SQLite](https://www.sqlite.org); no external software is needed.

- RSQLite is a DBI-compatible interface which means you primarily use functions defined in the DBI package, so you should always start by loading DBI, not RSQLite:

```{r}
library("DBI")
```

## Creating a new database

To create a new SQLite database, you simply supply the filename to `dbConnect()`:
    
```{r}
mydb <- dbConnect(RSQLite::SQLite(), "my-db.sqlite")
dbDisconnect(mydb)
unlink("my-db.sqlite")
```

If you just need a temporary database, use either `""` (for an on-disk database) or `":memory:"` or `"file::memory:"` (for a in-memory database). This database will be automatically deleted when you disconnect from it.

```{r}
mydb <- dbConnect(RSQLite::SQLite(), "")
dbDisconnect(mydb)
```

## Adding data to the database

1. Load a table using `read.csv()`

```{r}
airports <- read.csv("nycflights13/airports.csv")[, -1]
planes <- read.csv("nycflights13/planes.csv")[, -1]
```

2. Add to the database using `dbWriteTable()`:
    
```{r}
mydb <- dbConnect(RSQLite::SQLite(), "")
dbWriteTable(mydb, "airports", airports)
dbWriteTable(mydb, "planes", planes)
dbListTables(mydb)
```

## Queries

Queries in **RSQLite** pass SQL code directly, using `dbGetQuery()`:
    
```{r}
dbGetQuery(mydb, 'SELECT * FROM airports LIMIT 5')
```

```{r}
dbGetQuery(mydb, 'SELECT * FROM planes WHERE engines > 2')
```


## Batched queries

If you run a query and the results don't fit in memory, you can use `dbSendQuery()`, `dbFetch()` and `dbClearResults()` to retrieve the results in batches. By default `dbFetch()` will retrieve all available rows: use `n` to set the maximum number of rows to return.

```{r}
rs <- dbSendQuery(mydb, 'SELECT * FROM planes')
while (!dbHasCompleted(rs)) {
    df <- dbFetch(rs, n = 10)
    print(nrow(df))
}
dbClearResult(rs)
```

